# Reproducible Research: Peer Assessment 1
# Dakota Carter


## Loading and preprocessing the data
```{r, echo=TRUE}
library(plyr)
library(ggplot2)
library(knitr)
acdata <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
```{r, echo=TRUE}
sumbydate <- tapply(acdata$steps, acdata$date, FUN = sum, na.rm = TRUE)
mean(sumbydate, na.rm = T)
median(sumbydate, na.rm = T)

```

Here is the Histogram

```{r, echo=TRUE, fig.width=7, fig.height=7, fig.path='figures/'}
qplot(sumbydate, binwidth = 1000, xlab = "total number of steps taken each day")
```


## What is the highest average daily activity pattern?
```{r, echo=TRUE}
avgsteps <- aggregate(x = list(steps = acdata$steps), by = list(interval = acdata$interval), 
    FUN = mean, na.rm = TRUE)
avgsteps[which.max(avgsteps$steps), ]
```

Here is the Plot
```{r, echo=TRUE, fig.width=7, fig.height=7, fig.path='figures/'}
ggplot(data = avgsteps, aes(x = interval, y = steps)) + geom_line() + xlab("5-minute interval") + 
    ylab("average number of steps taken")
```



## Imputing missing values
To replace the values the following code will replace any missing value with the mean value of that 5
minute interval across all the days with data for that interval. 
```{r, echo=TRUE}
missing <- sum(is.na(acdata))
missing
```
There are 2304 rows with missing values
Below I replace the missing values with the average value for that time interval.

```{r, echo=TRUE}
imputemean <- function(x) replace(x, is.na(x), mean(x, na.rm = T))
newacdata <- ddply(acdata, ~ interval, transform, steps = imputemean(steps))
newsumbydate <- tapply(newacdata$steps, newacdata$date, sum)
newsumby <- as.vector(newsumbydate)
```

Here is the Histogram with the replaced values
```{r, echo=TRUE, fig.width=7, fig.height=7, fig.path='figures/'}
qplot(newsumbydate, binwidth = 1000, xlab = "total number of steps taken each day")
```

```{r, echo=TRUE}
mean(newsumby)
median(newsumby)
```

Imputing missing values gets rid of the scew! Yay!


## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
whichday <- function(date) {
        day <- weekdays(date)
        if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
                return("weekday") 
        else if (day %in% c("Saturday", "Sunday"))
                return("weekend")
        else stop("invalid date")
}
newacdata$date <- as.Date(newacdata$date)
newacdata$day <- sapply(newacdata$date, FUN = whichday)
avgsteps <- aggregate(steps ~ interval + day, data = newacdata, mean)
```

Here is the Panel Plot
```{r, echo=TRUE, fig.width=7, fig.height=7, fig.path='figures/'}
ggplot(avgsteps, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + 
    xlab("5-minute interval") + ylab("Number of steps")
```

The average weekend steps seem higher, especially late morning to evening. 
